---
title: "alex-flexdashboard"
author: "Hoang Phong Nguyen"
date: "5/21/2021"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
---

```{r libraries, echo=FALSE}
library(flexdashboard)
library(tidyverse)
library(naniar)
library(ggplot2)
library(kableExtra)
library(plotly)
```

```{r}
# read tidy data
fatalcrashes <- read.csv(here::here("data/cleaned_fatalcrashes.csv"))
fatalities <- read.csv(here::here("data/cleaned_fatalities.csv"))
```

Speed Limit Analysis {data-icon="fa-tachometer-alt"}
====================

```{r hp-clean-fatal}
fatalities$Speed.Limit <-
  str_replace(fatalities$Speed.Limit, "<40", "39")
  
fatal_sl <- fatalities %>% 
  mutate(Speed.Limit = as.numeric(Speed.Limit)) %>% 
  select(Crash.ID, Year, State, Speed.Limit) %>% 
  group_by(Year, Speed.Limit) %>% 
  count() %>% 
  pivot_wider(names_from = Speed.Limit,
              values_from = n) %>% 
  mutate("<= 40" = sum(`10`, `39`, `40`, `20`, `30`, `5`, `15`, `25`, 
                     na.rm = T)) %>% 
  mutate("70-90" = sum(`70`, `75`, `80`, `90`, na.rm = T)) %>% 
  mutate(">= 100" = sum(`100`, `110`, `130`, na.rm = T)) %>% 
  select(Year, `<= 40`, `50`, `60`, `70-90`, `>= 100`) # "NA")
```

```{r hp-clean-crash}
fatalcrashes$Speed.Limit <-
  str_replace(fatalcrashes$Speed.Limit, "<40", "39")
  
crash_sl <- fatalcrashes %>% 
  mutate(Speed.Limit = as.numeric(Speed.Limit)) %>% 
  select(Crash.ID, Year, State, Speed.Limit) %>% 
  group_by(Year, Speed.Limit) %>% 
  count() %>% 
  pivot_wider(names_from = Speed.Limit,
              values_from = n) %>% 
  mutate("<= 40" = sum(`10`, `39`, `40`, `20`, `30`, `5`, `15`, `25`, 
                     na.rm = T)) %>% 
  mutate("70-90" = sum(`70`, `75`, `80`, `90`, na.rm = T)) %>% 
  mutate(">= 100" = sum(`100`, `110`, `130`, na.rm = T)) %>% 
  select(Year, `<= 40`, `50`, `60`, `70-90`, `>= 100`) #, "NA")
```

Row
------------------

### Table 6.1: Yearly number of fatal crashes by speed limit

```{r hp-table}
crash_sl %>% kbl() %>% 
  kable_minimal()
```


```{r hp-pivot-longer}
crash_sl_l <- crash_sl %>% 
  pivot_longer(crash_sl, cols = -Year, 
             names_to = "Speed_limit", values_to = "n_crashes")

fatal_sl_l <- fatal_sl %>% 
  pivot_longer(fatal_sl, cols = -Year, 
             names_to = "Speed_limit", values_to = "n_fatalities")
```

### Figure 6.1: Comparing proportion of yearly fatal crash by speed limit

```{r hp-fig-year}
crash_sl_l %>% mutate(Speed_limit = fct_reorder(Speed_limit, n_crashes)) %>% 
  ggplot(aes(x = as.factor(Year), 
             y = n_crashes, fill = Speed_limit)) +
  geom_col(position = "fill") +
  scale_fill_brewer(palette="Set1") +
  theme(axis.title.x = element_blank())
ggplotly()
```

```{r hp-clean-compare}
comparison_crash_fatal <- 
  left_join(crash_sl_l, fatal_sl_l, 
            by = c("Year", "Speed_limit"))

comparison_pct <- comparison_crash_fatal %>%
  group_by(Speed_limit) %>% 
  summarise(n_crashes = sum(n_crashes),
         n_fatalities = sum(n_fatalities)) %>% 
  mutate("perc_fatalities" = (n_fatalities - n_crashes)*100/n_crashes) %>% 
  arrange(perc_fatalities)

comparison_l <- comparison_pct %>% 
  select(-perc_fatalities) %>% 
  pivot_longer(cols = c(n_crashes, n_fatalities), 
               names_to = "type", values_to = "n_fatal")
```

Row
--------------

### Figure 6.2: Comparing number of fatalities and fatal crashes by speed zone

```{r hp-fig-compare}
comparison_l %>% 
  mutate(Speed_limit = fct_reorder(Speed_limit, n_fatal)) %>%
  ggplot(aes(x = Speed_limit, 
             y = n_fatal, fill = type)) +
  geom_col(position = "dodge") +
  theme_bw() +
  labs(x = "Speed Limit (km/h)",
       y = "Count")
ggplotly()
```

### Figure 6.3: Percentage difference between fatalities and fatal crashes

```{r hp-fig-percent-diff}
comparison_pct %>% 
  mutate(Speed_limit = fct_reorder(Speed_limit, perc_fatalities)) %>%
  ggplot(aes(x = Speed_limit, y = perc_fatalities)) +
  geom_col(fill = "light blue") +
  theme_bw() +
  labs(x = "Speed Limit (km/h)",
       y = "Percent")
  ggplotly()
```

### Analysis

- Table 6.1 showed between 2010 and 2020, fatal crashes is reduced for higher speed zones but increased for lower speed zones

- The speed zone >= 100 km/h accounts for nearly 50% of all fatal crashes in Australia

- Number of fatal crashes and fatalities both increased with speed limits

- Fatalities is 3% higher than number of fatal crashes at lower speed zone, but increase up to 13% for higher speed zone
