---
title: "Group Project"
author: "T11_Wed_reprex"
output: 
  bookdown::html_document2: default
  bookdown::pdf_document2: default
  fig_caption: yes
  toc: yes
keep_md: TRUE
editor_options: 
  chunk_output_type: console
---


# Data Description
The data we have chosen for this project is the Australian Road Deaths Database (ARDD) provided by the Australian Bureau of Infrastructure and Transport Research Economics (BITRE). The records are updated monthly from 1986 to 2021. Here is the link to the database : https://data.gov.au/dataset/ds-dga-5b530fb8-526e-4fbf-b0f6-aa24e84e4277/details. 

There are two data files. The fatality data contains the demographic and details of people who have died within 30 days of the traffic accident due to injuries caused by an Australia road crash. The fatal crashes data contains the records for the crash, including information like the road type, and speed limit. The two data files can be connected by the crash ID. See Appendix for more detailed descriptions of each variable. 


# Research Questions:
Q1: What demographic has a higher rate of traffic accidents?

Q2: Ranking of vehicle types by number of accidents

Q3: How does the car accident fatality link with the district and accident type? 

Q4: Is there any correlation between accidents and holiday? And why are some explanations for the relation(s) or lack thereof.

Q5 :Has accident rate decreased or increased within the past decade (2010- 2020)? 

Q6: Possible correlations between number of fatalities and speed limit zone 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r libraries, echo=FALSE}
# Loading package tidyverse:
library(tidyverse)
library(naniar)
library(ggplot2)
library(kableExtra)
```


```{r}
# read tidy data
fatalcrashes <- read.csv(here::here("data/cleaned_fatalcrashes.csv"))
fatalities <- read.csv(here::here("data/cleaned_fatalities.csv"))
```

# Data Analysis

##Q1:


##Q2:


##Q3:


##Q4:

```{r}
holidays <- fatalcrashes %>%
  select(State, Year, Month, Christmas.Period, Easter.Period, Time.of.day)
```

```{r}
# detect the Christmas and Easter period month first
holidays1 <- holidays %>%
  select(Month, Christmas.Period, Easter.Period) %>%
  filter(Christmas.Period == "Yes" | Easter.Period == "Yes")

unique(holidays1$Month)
```

+ The month shows that the Christmas period is in December and January. And the Easter period is in March and April.

```{r}
library(lubridate)
library(plotly)
# filter month and check the holidays
monthly <- holidays %>%
  mutate(Month = as.factor(Month)) %>%
  group_by(Month, State)%>%
  summarise(count = n())


#draw line plot
p1 <- ggplot(monthly, aes(x = Month, 
                    y = count, 
                    group = State, 
                    color = State)) +
  geom_line()+
  geom_point(alpha = 0.8, size = 2)+
  scale_color_brewer(palette = "Set1")+
  labs(title = "Fatal crashes happens monthly in different states")+
  theme_light()
ggplotly(p1)


```


+ The month shows that the Christmas period is in December and January. And the Easter period is in March and April.

+ Only NSW  and NT has little increase during Easter period, while other states got a decline during such period.

+ For Christmas period, the number of crashes only increases in NSW, other states decrease or stay the same.

+ There is no clear correlation between fatal crashes with Christmas or Easter holidays.


```{r}
# check the correlation with night and day
night <- holidays %>%
  select(Year, Time.of.day, State) %>%
  mutate(Time.of.day = as.factor(Time.of.day),
         Year = as.factor(Year)) %>%
  group_by(State, Time.of.day, Year) %>%
  count()

# plot the day and night by year
p3 <- ggplot(night, 
       aes(x = n, y = Year, fill = Time.of.day))+
  geom_bar(stat = "identity", position = "fill")+
  facet_wrap(~State, nrow = 1)+
  geom_vline(xintercept = 0.5, color = "yellow")+
  scale_fill_manual(values=c("#ae7181", "#350631"))+
  labs(title = "Fatal crashes happens by day and night")+
  theme_classic()
  
ggplotly(p3)
```

+ Obviously, there are more crashes happen on the days than night.

+ There is more difference in ACT.

+ In NT, some crashes happens more at night than day in 2020, 2018, 2015.

##Q5:


##Q6:
