---
title: "Group Project"
author: "T11_Wed_reprex"
output: 
  bookdown::pdf_document2: default
  bookdown::html_document2: default
  fig_caption: yes
  toc: yes
keep_md: TRUE
editor_options: 
  chunk_output_type: console
---


# Data Description
The data we have chosen for this project is the Australian Road Deaths Database (ARDD) provided by the Australian Bureau of Infrastructure and Transport Research Economics (BITRE). The records are updated monthly from 1986 to 2021. Here is the link to the database : https://data.gov.au/dataset/ds-dga-5b530fb8-526e-4fbf-b0f6-aa24e84e4277/details. 

There are two data files. The fatality data contains the demographic and details of people who have died within 30 days of the traffic accident due to injuries caused by an Australia road crash. The fatal crashes data contains the records for the crash, including information like the road type, and speed limit. The two data files can be connected by the crash ID. See Appendix for more detailed descriptions of each variable. 


# Research Questions:
Q1: What demographic has a higher rate of traffic accidents?

Q2: Ranking of vehicle types by number of accidents

Q3: How does the car accident fatality link with the district and accident type? 

Q4: Is there any correlation between accidents and holiday? And why are some explanations for the relation(s) or lack thereof.

Q5 :Has accident rate decreased or increased within the past decade (2010- 2020)? 

Q6: Possible correlations between number of fatalities and speed limit zone 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r libraries, echo=FALSE}
# Loading package tidyverse:
#install.packages("DT")
library(tidyverse)
library(naniar)
library(ggplot2)
library(kableExtra)
<<<<<<< HEAD
library(plotly)
=======
library(flexdashboard)
library(DT)
>>>>>>> eb19d8083d23fe33f9af1e3393ac41f1a8163247
```


```{r}
# read tidy data
fatalcrashes <- read.csv(here::here("data/cleaned_fatalcrashes.csv"))
fatalities <- read.csv(here::here("data/cleaned_fatalities.csv"))
head(fatalcrashes)
head(fatalities)
```

# Data Analysis

## Q1:What demographic has a higher rate of road accidents or fatal crashes?
+ sex as a factor in accident
+ Age as a factor in accident

```{r setup}
library(tidyverse)
library(ggplot2)
library(readr)
```

```{r Reading-data}
fatalities_new <- read.csv(here::here("data/cleaned_fatalities.csv"))
```

```{r setting-data}

accidents <- fatalities_new %>% 
  filter(!is.na(Gender)) %>% 
  select( "Gender", "Age.Group") %>% 
  group_by( Age.Group) %>% 
  count(Gender, Age.Group) %>% 
  rename(`Total no. of accidents` = n)

head(accidents)
```


```{r Graph1, Q1Fig1, fig.cap= "Australian road deaths database and fatal crashes, by age and gender", fig.show="hold" }
graph1 <- ggplot(data = accidents, aes(x = Age.Group,
                             y = `Total no. of accidents`, 
                             fill = Gender)) +
  geom_bar(stat="identity", position = "dodge") +
  scale_fill_brewer(palette = "Pastel2")+
  theme_classic() +
  xlab("Age Group") +
  ylab("Total number of Accidents") 


graph1
``` 

+ The Australian Road Deaths Database provides basic details of road transport crash fatalities in Australia as reported by the police each month to the State and Territory road safety authorities.
+ Road deaths from recent months are preliminary and the series is subject to revision.
+ The above graph \@ref(fig:Q1Fig1) explains, what demographic has a higher rate of traffic accidents considering **Age.Group** and **Gender** as a factor.
+ The age group between **40 to 64** in both the genders, had the maximum risk of fatal crash involvements compared to the rest of the age groups. 
+ **Male** drivers compared to **Female** for the same age group have the highest rate of road accidents i.e. more than **3000** fatal crashes. Hence, Male drivers between age group **40 t0 64** are more susceptible to the fatal crash involvement.    
 



## Q2:Which road user is more prone to accident over the year?

```{r, setting-data-roaduser}
accidents1 <- fatalities_new %>% 
  mutate(Year = as.factor(Year)) %>% 
  select("Year", "Road.User") %>% 
  filter(Road.User %in% c("Driver", "Motorcycle pillion passenger", "Motorcycle rider", "Passenger", "Pedal cyclist", "Pedestrian")) %>% 
  group_by(Road.User, Year) %>% 
  count(Road.User) %>% 
  rename(`Accident Total count` = n)

head(accidents1)

```


```{r Graph2, Q2Fig2, fig.cap= "Australian road deaths database and fatal crashes, by age and gender", fig.show="hold" ,fig.width=5}
graph2 <- ggplot(accidents1, aes(x = Year, 
                 y = `Accident Total count`)) +
  geom_jitter(position = position_jitter(width = 0.5), aes(color = factor(Road.User)), size = 1.5, alpha = 5) +
  facet_wrap(Road.User~.) +
  theme(axis.text.x = element_text(size =10)) +
  theme_dark() +
  xlab("Year") +
  ylab("Total number of Accidents")

graph2

```

+ This analysis is with respect to the vulnerable Road Users who are prone to road accidents over the years in Australia. 
+ According the above graph \@ref(fig:Q2Fig2), **Driver** is one of the most vulnerable road user prone to accident through-out the years.
+ In **2010** and **2016** driver experienced the peak where there were maximum number of casualties. 
+ Along with that even **Pedestrian**, **Passenger** and **Motorcycle rider** are among the second most vulnerable road user, prone to accident. 
+ The number of casualties for **Motorcycle rider** and **Pedestrian** has mostly been constant through-out the years from **2010 till 2020** 
+ The casualties for **Passenger** shows a slight decrease over the years. 



##Q3:

Question 3

#How does the car accident fatality link with the district and accident type? 



First of all, we group the data by state and the type of car which have involved in the fatal accidents.
```{r, eval=TRUE, echo=FALSE}

#no. of fatal accidents with bus involved
bus_involved <- fatalcrashes %>%
  filter(Bus.Involvement == "Yes") %>%
  group_by(State)%>%
  count() %>%
  rename("Bus_Involved" = n) 

#no. of fatal accidents with heavy truck involved
heavytruck_involved <- fatalcrashes %>%
  filter(Heavy.Rigid.Truck.Involvement == "Yes") %>%
  group_by(State)%>%
  count()%>%
  rename("Heavytruck_Involved" = n)

#no. of fatal accidents with articulated truck involved
articulated_truck_involved <- fatalcrashes %>%
  filter(Articulated.Truck.Involvement == "Yes") %>%
  group_by(State)%>%
  count()%>%
  rename("Articulated_truck_Involved" = n)

head(bus_involved)
head(heavytruck_involved)
head(articulated_truck_involved)
```


```{r, eval=TRUE, echo=FALSE}
#join the above 3 tables and turns them to long format
accident_car_type_count <- bus_involved %>%
  full_join(heavytruck_involved) %>%
  full_join(articulated_truck_involved) %>%
  pivot_longer(cols = c("Bus_Involved", "Heavytruck_Involved", "Articulated_truck_Involved"),  
               names_to = "Type",             
               values_to = "Count")
head(accident_car_type_count)
```
In this table, we have got the count of different car type of fatal accidents happened group by different states. This data is in long format which will be easier for us to plot the bar chart for better visualization. 


```{r, eval=TRUE, echo=FALSE}
#plot the bar chart for the statistic of accident car type in different states
ggplot(accident_car_type_count, 
       aes( x = State, y = Count, fill = Type)) + 
  geom_col(position = "dodge") + 
  scale_fill_brewer( palette = "Set1" )+
  ggtitle("Number of Fatal Accidents with \nDifferent Car Type in Different State")

```

From the graph above:
 -NSW is the state with highest number of fatal accidents
 -follow by Qld and Vic. 
 
In all the states:
 -more than half of the total fatal accidents with articulated truck involved. 
 -Accidents with Heavy Truck involved account for the second large portion of the total accidents
 -follow by bus involved accidents. 
 
Interpretation:
Articulated truck is the heaviest among these 3 type of cars. Heavier car impose a greater impact on any accident which cause a higher fatalities. 

Articulated truck is a vehicle which has a permanent or semi-permanent pivot joint in its construction. Only large vehicles need this structure which is easier for them to turn. But it also impose an additional risk for higher fatalities in an accident due to the complex mechanical structure.

Then, we group the data by states and the type of Road Users which have involved in the fatal accidents.


```{r, eval=TRUE, echo=FALSE}
#How does the car accident fatality link with the district and accident type? 
accident_road_user_count <- fatalities %>%
  group_by(State, Road.User) %>%
  count() %>%
  rename(Count = n)
head(accident_road_user_count)
```


```{r,eval=TRUE, echo=FALSE}
#plot the bar chart for the statistic of accident road user type in different states
ggplot(accident_road_user_count, 
       aes( x = State, y = Count, fill = Road.User)) + 
  geom_col(position = "dodge") +
  scale_fill_brewer( palette = "Set1" )+
  labs(title = "Number of Fatal Accidents \nwith Different Road User in Different State")
```

From the graph above:
 
In all the states:
 -more than half of the total fatal accidents with drivers involved. 
 -Accidents with Passenger involved account for the second large portion of the total accidents
 -follow by motorcycle rider and then pedestrian involved accidents. 
 
Interpretation:
Most fatal accidents with the drivers and passengers involved. relatively less pedestrian involved in fatal car accidents. We may interpret that most of the people drive in Australia, that is the reason more drivers and passenger involved in fatal car accidents. Another interpretation is most of the car accidents happened in highway or it is a crashed between cars or crashed between car and the road.


Then, we group the data by the crash type and count the number of it.


```{r, eval=TRUE, echo=FALSE}
crash_type <- fatalities %>%
  group_by(State, Crash.Type) %>%
  count() %>%
  rename(Count = n)
head(crash_type)
```

```{r, eval=TRUE, echo=FALSE}
#plot the bar chart for the statistic of accident road user type in different states
ggplot(crash_type, 
       aes( x = State, y = Count, fill = Crash.Type)) + 
  geom_col(position = "dodge") +
  scale_fill_brewer( palette = "Set1" )+
  labs(title = "Number of Fatal Accidents \nwith Different Crash Type in Different State")
```
From the graph above:
 
In all the states except ACT and Tas:
 -all of the states have more single car crash than multiple car crash in terms of fatal car crashes.


How does the car accident fatality link with the district?
```{r, eval=TRUE, echo=FALSE}
#no. of fatal accidents in different district **with** NA value and arrange in decending order
district_fatalities <- fatalities %>%
  group_by(State,National.LGA.Name.2017) %>%
  count()%>%
  arrange(desc(n))
head(district_fatalities)

#no. of fatal accidents in different district **without** NA value and arrange in decending order
district_fatalities_without_NA <- fatalities %>%
  group_by(State,National.LGA.Name.2017) %>%
  count()%>%
  filter(!is.na(National.LGA.Name.2017))%>%
  arrange(desc(n))
head(district_fatalities_without_NA)
```

From the table above:
-first table contain value with NA
  -in this table, we don't have detail district information for some cases.              -However, in general, we could see that QLD, NSW, and Vic, these 3 states are top 3 in terms of number of fatal accidents. 
Be aware, this count do not reflect the actual ranking for the number of accidents.
-second table contain value without NA
  -in this table, we can conclude:
  +Brisbane in Qld is the district which has the most fatal car accidents happened, which has 87 cases within this 10 years.
  +Follow by Central Coast in NSW which is 82 fatal car accidents happened
  +The third one is Gold Coast which has 68 cases in this 10 years.
  
Interpretation:
QLD, NSW, and Vic are the states that contain top 3 largest amount of population. Therefore, there will be higher chance of having a car accidents. 
It is the same as the district. Brisbane, Central Coast and Gold Coast are the districts with higher population which may leads to higher chance of an accident happened.

##Q4:

```{r}
holidays <- fatalcrashes %>%
  select(State, Year, Month, Christmas.Period, Easter.Period, Time.of.day)
```

```{r}
# detect the Christmas and Easter period month first
holidays1 <- holidays %>%
  select(Year, Month, Christmas.Period, Easter.Period) %>%
  filter(Christmas.Period == "Yes" | Easter.Period == "Yes")

unique(holidays1$Month)
```

+ The month shows that the Christmas period is in December and January. And the Easter period is in March and April.

```{r}
# filter month and check the holidays
monthly <- holidays %>%
  mutate(Month = as.factor(Month)) %>%
  group_by(Month, State)%>%
  count()
```

```{r Christmas, fig.align='center', fig.cap= " The fatal crashes happen in Christmas period from 2018 to 2020"}
# compare the period of holidays in the last 3 years
library(plotly)
period <- holidays %>%
  filter(Year %in% c("2018" : "2020"),
         Month %in% c ("11", "12", "1", "2", "3", "4", "5")) %>%
  mutate(Month = as.factor(Month)) %>%
  group_by(Year, Month, State) %>%
  count() 

period$Month <- factor(period$Month, levels = c("11", "12", "1", "2", "3", "4", "5"))
orderper <- period[order(period$Month),]


# filter Christmas data
Christmas <- orderper %>%
  filter(Month %in% c("11", "12", "1", "2"))

## plot the Christmas change  of recent 6 years
Christmas_fig <- ggplot(Christmas, aes(x = Month, y = n, color = State, group = State))+
  geom_line()+
  geom_point(alpha = 0.8, size = 2)+
  scale_color_brewer(palette = "Set1")+
  theme_light()+
  facet_wrap(~Year)
ggplotly(Christmas_fig)
```

It can be seen from \@ref(fig:Christmas) that :

+ The number of crashes increases in VIC, WA and NSW in 2018 and 2019 respectively during Christmas period.

+ The OLD got a sudden decline in Christmas period of 2020, maybe due to the effectof COVID virus.

+ For other states, the crashes either fluctuated or drop. Therefore, there is no dominate correlation between crashes and Christmas.



```{r Easter, fig.align='center', fig.cap= "The crashes in Easter period from 2018 to 2020"}
# filter Easter data
Easter <- orderper %>%
  filter(Month %in% c("2", "3", "4", "5"))

## plot the Christmas change  of recent 6 years
Easter_fig <- ggplot(Easter, aes(x = Month, y = n, color = State, group = State))+
  geom_line()+
  geom_point(alpha = 0.8, size = 2)+
  scale_color_brewer(palette = "Set1")+
  theme_classic()+
  facet_wrap(~Year)
ggplotly(Easter_fig)
```

In Figure \@ref(fig:Easter), in 2018, only WA has a slightly increase in Easter period. For Victoria, the number of crashes drops dramatically in March to April of 2018 and 2020. While other states fluctuate in these period. There is no clear correlation between cashes in Easter period in different states.


```{r night, fig.align='center', fig.cap= "Fatal crashes happens by day and night"}
# check the correlation with night and day
night <- holidays %>%
  select(Year, Time.of.day, State) %>%
  mutate(Time.of.day = as.factor(Time.of.day),
         Year = as.factor(Year)) %>%
  group_by(State, Time.of.day, Year) %>%
  count()

# plot the day and night by year
p3 <- ggplot(night, 
       aes(x = n, y = Year, fill = Time.of.day))+
  geom_bar(stat = "identity", position = "fill")+
  facet_wrap(~State, nrow = 1)+
  geom_vline(xintercept = 0.5, color = "yellow")+
  scale_fill_manual(values=c("#ae7181", "#350631"))+
  labs(title = "Fatal crashes happens by day and night")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90))
  
ggplotly(p3)
```

The Figure \@ref(fig:night) compares the fatal crashes that happen in days and night.
+ Obviously, there are more crashes happen on the days than night.

+ In 2018 and 2013, the crashes of ACT happen more on night rather than day.

+ In NT, some crashes happens more at night than day in 2020, 2018, 2015.




##Q5:

```{r accident data}
accident <- fatalcrashes %>%
  mutate(Month = as.factor(Month),
         Year = as.factor(Year),
         date = paste(Month, Year, sep = "/")) %>%
  select(c("State", "Month", "Year", "date"))
count(accident, Year) %>% kable(bookdown = TRUE)
```
Although 2019 was the year the pandemic started, it surprisingly was not the year with the lowest number of fatal crashes in total. 
```{r accident aggregatted across years}
accident2 <-  accident %>%
  group_by(date) %>%
  add_tally() %>%
  select(-State) %>%
  distinct()

ggplot(accident2 %>%
         mutate(Month = as.numeric(Month)),
       aes(x = Month,
           y = n,
           colour= Year)) +
  facet_wrap(~Year) +
  geom_line() + geom_point()+
  scale_colour_brewer(palette = "Set3")+
  theme_minimal()+
  ylab("Number of Crashes") +
  scale_x_continuous(breaks=seq(0,12,1))
```

```{r accident time series by state, fig.width= 10}
accident3 <- accident %>%
        group_by(State, date) %>%
        add_tally()%>%
        distinct() %>%
        arrange(State, date)
        
ggplot(accident3,
       aes(x = Month,
           y = n,
           fill = State))+
  geom_col() +
  facet_grid(State + .~Year)+ 
  scale_fill_brewer(palette = "Set1")+
  theme_minimal()
```

##Q6:

```{r hp-clean-fatal, echo=F}
fatalities$Speed.Limit <-
  str_replace(fatalities$Speed.Limit, "<40", "39")
  
fatal_sl <- fatalities %>% 
  mutate(Speed.Limit = as.numeric(Speed.Limit)) %>% 
  select(Crash.ID, Year, State, Speed.Limit) %>% 
  group_by(Year, Speed.Limit) %>% 
  count() %>% 
  pivot_wider(names_from = Speed.Limit,
              values_from = n) %>% 
  mutate("<= 40" = sum(`10`, `39`, `40`, `20`, `30`, `5`, `15`, `25`, 
                     na.rm = T)) %>% 
  mutate("70-90" = sum(`70`, `75`, `80`, `90`, na.rm = T)) %>% 
  mutate(">= 100" = sum(`100`, `110`, `130`, na.rm = T)) %>% 
  select(Year, `<= 40`, `50`, `60`, `70-90`, `>= 100`) # "NA")
```

```{r hp-clean-crash, echo=F}
fatalcrashes$Speed.Limit <-
  str_replace(fatalcrashes$Speed.Limit, "<40", "39")
  
crash_sl <- fatalcrashes %>% 
  mutate(Speed.Limit = as.numeric(Speed.Limit)) %>% 
  select(Crash.ID, Year, State, Speed.Limit) %>% 
  group_by(Year, Speed.Limit) %>% 
  count() %>% 
  pivot_wider(names_from = Speed.Limit,
              values_from = n) %>% 
  mutate("<= 40" = sum(`10`, `39`, `40`, `20`, `30`, `5`, `15`, `25`, 
                     na.rm = T)) %>% 
  mutate("70-90" = sum(`70`, `75`, `80`, `90`, na.rm = T)) %>% 
  mutate(">= 100" = sum(`100`, `110`, `130`, na.rm = T)) %>% 
  select(Year, `<= 40`, `50`, `60`, `70-90`, `>= 100`) #, "NA")
```


```{r hp-table}
crash_sl %>% kbl(caption = "Yearly number of fatal crashes by speed limit (km/h)") %>% 
  kable_minimal()
```

### Analysis

The speed limits are split into 5 respective parts which is 40 and below, 50, 60, 70-90 and 100 and above, all measured in kilometer per hour (km/h). Areas which have less than 40 km/h are often shared zones, school zones and places with high density of pedestrian. 50 km/h are default speed limit within built up areas in every state in Australia except for Northern Territory. 60 km/h are sub-arterial roads, as well as the default speed within built up area in Northern Territory. 70-90 are connector and small highways. 100 and above area highways speed limits.

Between 2010 and 2020, table \@ref(tab:hp-table) showed that the overall trend in fatal crashes is reduced for speed limit zones outside built-up areas but increased for zones within built-up areas. However, within the period of 2015-2016, all zones see an increase in number of fatal crashes. The number of fatal crashes reduced or increased in varying number as well as rate across the speed zones. With the speed zones of 100 and above and 60 km/h decreased by largest proportion, about 25%.

```{r hp-pivot-longer, echo=F}
crash_sl_l <- crash_sl %>% 
  pivot_longer(crash_sl, cols = -Year, 
             names_to = "Speed_limit", values_to = "n_crashes")

fatal_sl_l <- fatal_sl %>% 
  pivot_longer(fatal_sl, cols = -Year, 
             names_to = "Speed_limit", values_to = "n_fatalities")
```


```{r hp-fig-year, fig.cap="Comparing proportion of yearly fatal crash by speed limit"}
crash_sl_l %>% mutate(Speed_limit = fct_reorder(Speed_limit, n_crashes)) %>% 
  ggplot(aes(x = as.factor(Year), 
             y = n_crashes, fill = Speed_limit)) +
  geom_col(position = "fill") +
  scale_fill_brewer(palette="Set1") +
  theme(axis.title.x = element_blank())
ggplotly()
```

### Analysis

One notable point as figure \@ref(fig:hp-fig-year) showed is throughout the period analyzed, not only is the speed zone of 100 km/h and above has highest percentage of fatal crash, but it accounts for nearly 50% of all fatal crashes in Australia. While 50 km/h and below speed zones account for less than 10% of all fatal crashes. It is generally well known that the higher the speed a vehicle is traveling at, the longer the stopping distance is, and thus fatal crash is more likely to occur on highways. Kloeden, Woolley & McLean (2007) found that following the reduction of urban speed limit from 60 to 50 km/h in South Australia in 2003, there was a 23% reduction in fatal crashes in 50 km/h zones and 16% reduction in 60 km/h zones.

```{r hp-clean-compare}
comparison_crash_fatal <- 
  left_join(crash_sl_l, fatal_sl_l, 
            by = c("Year", "Speed_limit"))

comparison_pct <- comparison_crash_fatal %>%
  group_by(Speed_limit) %>% 
  summarise(n_crashes = sum(n_crashes),
         n_fatalities = sum(n_fatalities)) %>% 
  mutate("perc_fatalities" = (n_fatalities - n_crashes)*100/n_crashes) %>% 
  arrange(perc_fatalities)

comparison_l <- comparison_pct %>% 
  select(-perc_fatalities) %>% 
  pivot_longer(cols = c(n_crashes, n_fatalities), 
               names_to = "type", values_to = "n_fatal")
```


```{r hp-fig-compare, fig.cap="Comparing number of fatalities and fatal crashes by speed zone"}
comparison_l %>% 
  mutate(Speed_limit = fct_reorder(Speed_limit, n_fatal)) %>%
  ggplot(aes(x = Speed_limit, 
             y = n_fatal, fill = type)) +
  geom_col(position = "dodge") +
  theme_bw() +
  labs(x = "Speed Limit (km/h)",
       y = "Count")
ggplotly()
```



```{r hp-fig-percent-diff, fig.cap="Percentage difference between fatalities and fatal crashes"}
comparison_pct %>% 
  mutate(Speed_limit = fct_reorder(Speed_limit, perc_fatalities)) %>%
  ggplot(aes(x = Speed_limit, y = perc_fatalities)) +
  geom_col(fill = "light blue") +
  theme_bw() +
  labs(x = "Speed Limit (km/h)",
       y = "Percent")
  ggplotly()
```

### Analysis

Figure \@ref(fig:hp-fig-compare) illustrated that not only the number of fatal crashes increased with speed limits, but number of fatalities increased as well, and in higher proportion than the increase of crashes. The difference between number of fatalities and number of crashes are made clearer in figure \@ref(fig:hp-fig-percent-diff). The percentage difference is similar in speed zones of 60 km/h and below, at around 3%. However, the difference for 70-90 km/h is approx. 8% and for highway speed is nearly 13%.

### Citations

Kloeden, C., Woolley, J., & McLean, A. J. (2007, October). A follow-up evaluation of the 50km/h default urban speed limit in South Australia. In Proceedings of.