---
title: "HR Movement dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    runtime: shiny
    source_code: embed
---

```{r setup, include=FALSE}
# install.packages(c("tidyverse", "readxl", "plotly", "ggmap"))
library(flexdashboard)
library(ggplot2)
library(readxl)
library(tidyverse)
library(plotly)
library(ggmap)
```

```{r load data}
fatalcrashes <- read.csv(here::here("data/cleaned_fatalcrashes.csv"))
fatalities <- read.csv(here::here("data/cleaned_fatalities.csv"))
```

Question 5 {data-icon="fa-user-plus"}
==================
Column
--------

### Number of Car Accidents per Month from 2010-2020 by State
  - ACT has the lowest number of accidents across the years while NSW has the highest number of cases on average.
  - there doesn't seem to be any pattern across the years.
```{r}
accident <- fatalcrashes %>%
  mutate(Month = as.factor(Month),
         Year = as.factor(Year),
         date = paste(Month, Year, sep = "/")) %>%
  select(c("State", "Month", "Year", "date"))

accident2 <-  accident %>%
  group_by(date) %>%
  add_tally() %>%
  select(-State) %>%
  distinct()

plot5_1 <- ggplot(accident2 %>%
         mutate(Month = as.numeric(Month)),
       aes(x = Month,
           y = n,
           colour= Year)) +
  facet_wrap(~Year) +
  geom_line() + geom_point()+
  scale_colour_brewer(palette = "Set3")+
  theme_minimal()+
  ylab("Number of Crashes") +
  scale_x_continuous(breaks=seq(0,12,1))

ggplotly(plot5_1)
```

### Number of Car Accidents per Month from 2010-2020 by State
  - ACT has the lowest number of accidents across the years while NSW has the highest number of cases on average.
  - there doesn't seem to be any pattern across the years.
```{r}
accident3 <- accident %>%
        group_by(State, date) %>%
        add_tally()%>%
        distinct() %>%
        arrange(State, date)
        
plot5_2 <- ggplot(accident3,
       aes(x = Month,
           y = n,
           fill = State))+
  geom_col() +
  facet_grid(State + .~Year)+ 
  scale_fill_brewer(palette = "Set1")+
  theme_minimal()

ggplotly(plot5_2)
```

